---
title: "datelife_class"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Resources:

collaborative document: https://pad.carpentries.org/datelife 
datelife documentation: http://phylotastic.org/datelife/index.html

## Installing datelife

```{r}
devtools::install_github("phylotastic/datelife")
```

### Get a list of taxonomic names that you are interested on, it can/t be common names

- We can get tip labels from a tree
- We can create  a list of species or higher group names
- to standardize to the taxonomy, we use `make_datelife_query`

Case when we have a single taxon name
```{r}
my_group <- "fringillidae"
```

Case when we want names from a tree

```{r}
my_tree <- ape::read.tree(file = "data/krenek_etal_2013.tre")
my_tree
str(my_tree)
my_tree$tip.label

# get species names
my_group <- my_tree$tip.label


my_group <- gsub("'", "", my_group)

my_group_list <- strsplit(my_group, split = " ")

my_group <- sapply(my_group_list, function(x) paste0(x[1:2], collapse = "_"))  
```

```{r}
my_query <- datelife::make_datelife_query(input = my_group)

my_query <- datelife::make_datelife_query("mantodea", get_spp_from_taxon = TRUE)
str(my_query)
# check how many taxon names you got:
length(my_query$cleaned_names)
```



